### 2022-06-29 Corr.tex
Javier changes 2022-06-28 evening

\begin{table*}
\begin{center}
\caption{Linear regressions values in the form Y = aX + b between our turbulent parameters obtained using the chi-square statistic and properties of each region (Table \ref{tab:regions-properties}). The fifth column, $r$, is the Pearson correlation coefficient and the last column is the $p$-value. This results were obtained using the procedure in \citet{2007ApJ...665.1489K}.}
\begin{tabular}{cccccc}
\hline
            $Y$ &                   $X$ &                 $a$ &                 $b$ &       $r$ &      $p$ \\
\hline
      log $r_0$ &         log $D$ &   0.94 $\pm$ 0.33 &  -1.66 $\pm$ 0.72 &   0.86 &  0.003 \\
   log \(\sigma\pos\) &        log L(H) &    0.28 $\pm$ 0.1 &  -9.86 $\pm$ 3.71 &   0.86 &  0.003 \\
 \(\sigma\pos\) &  $\sigma_{\text{los}}$ &   1.03 $\pm$ 0.44 &   7.35 $\pm$ 2.83 &   0.78 &   0.01 \\
   log \(\sigma\pos\) &         log $D$ &   0.26 $\pm$ 0.18 &   0.19 $\pm$ 0.39 &   0.64 &   0.06 \\
        log $m$ &        log distance &  -0.02 $\pm$ 0.04 &   0.04 $\pm$ 0.08 &   -0.4 &   0.29 \\
       \(\sigma\pos\) &         $r_{0}$ &    0.4 $\pm$ 0.69 &   4.61 $\pm$ 3.25 &   0.37 &  0.329 \\
            $m$ &        \(\sigma\pos\) &  -0.01 $\pm$ 0.03 &   1.12 $\pm$ 0.21 &  -0.37 &  0.331 \\
            $m$ &         $r_{0}$ &   0.02 $\pm$ 0.03 &   0.96 $\pm$ 0.16 &  -0.09 &  0.824 \\
\bottomrule
\end{tabular}\label{tab:RestStats}
\end{center}
\end{table*}

%<10\(^{-5}\)