### 2022-06-29 Corr.tex
Improve formatting of Table 4

\begin{table*}
\begin{center}
\caption{Linear regressions values in the form Y = aX + b between our turbulent parameters obtained using the chi-square statistic and properties of each region (Table \ref{tab:regions-properties}). The fifth column, $r$, is the Pearson correlation coefficient and the last column is the $p$-value. This results were obtained using the procedure in \citet{2007ApJ...665.1489K}.}
\begin{tabular}{RRRRRR}
  \toprule
  Y &                   X &                 a &                 b &       r &      p \\
  \midrule
  \log r_0 &         \log D_{\hii} &   0.94 \pm 0.33 &  -1.66 \pm 0.72 &   0.86 &  \mathbf{0.003} \\
  \log \sigma\pos &        \log L(\ha) &    0.28 \pm 0.10 &  -9.86 \pm 3.71 &   0.86 &  \mathbf{0.003} \\
  \sigma\pos &  \sigma_{\text{los}} &   1.03 \pm 0.44 &   7.35 \pm 2.83 &   0.78 &   \mathbf{0.010} \\[\smallskipamount]
  \log \sigma\pos &         \log D_{\hii} &   0.26 \pm 0.18 &   0.19 \pm 0.39 &   0.64 &   0.060 \\
  \log m &        \log d &  -0.02 \pm 0.04 &   0.04 \pm 0.08 &   -0.40 &   0.290 \\
  \sigma\pos &         r_{0} &    0.40 \pm 0.69 &   4.61 \pm 3.25 &   0.37 &  0.329 \\
  m &        \sigma\pos &  -0.01 \pm 0.03 &   1.12 \pm 0.21 &  -0.37 &  0.331 \\
  m &         r_{0} &   0.02 \pm 0.03 &   0.96 \pm  0.16 &  -0.09 &  0.824 \\
  \bottomrule
\end{tabular}\label{tab:RestStats}
\end{center}
\end{table*}

%<10\(^{-5}\)

%%% Local Variables:
%%% mode: latex
%%% TeX-master: strucfunc_paper
%%% End: