### 2022-09-20 Corr.tex
Javier changes 2022-09-20

\begin{table*}
\begin{center}
\caption{Linear regressions values in the form Y = aX + b between our turbulent parameters obtained using the chi-square statistic and properties of each region (Table \ref{tab:regions-properties}). The fifth column, $r$, is the Pearson correlation coefficient and the last column is the $p$-value. This results were obtained using the procedure in \citet{2007ApJ...665.1489K}.}
\begin{tabular}{RRRRRR}
  \toprule
  Y &                   X &                 a &                 b &       r &      p \\
  \midrule
  \log r_0 &         \log D_{\hii} &   0.95 \pm 0.33 &  -1.68 \pm 0.71 &   0.86 &  \mathbf{0.003} \\
  \log \sigma\pos &        \log L(\ha) &    0.25 \pm 0.11 &  -9.01 \pm 4.26 &   0.81 &  \mathbf{0.008} \\
  \sigma\los &  \sigma\pos &   1.03 \pm 0.45 &   7.40 \pm 2.88 &   0.78 &   \mathbf{0.010} \\[\smallskipamount]
  \log \sigma\pos &         \log D_{\hii} &   0.26 \pm 0.18 &   0.20 \pm 0.39 &   0.64 &   0.06 \\
  \log \sigma\pos &   \log r_{0} &    0.19 \pm 0.18 &   0.68 \pm 0.13 &   0.51 &  0.16 \\
  \log m &  \log d &  -0.02 \pm 0.04 &   0.04 \pm 0.08 &   -0.40 &   0.29 \\
  \log m &  \log  \sigma\pos &  -0.11 \pm 0.20 &   0.09 \pm 0.16 &  -0.37 &  0.33 \\
  \log m &  \log  r_{0} &   -0.02 \pm 0.07 &   0.02 \pm  0.05 &  -0.28 &  0.47 \\
  \bottomrule
\end{tabular}\label{tab:RestStats}
\end{center}
\end{table*}

%<10\(^{-5}\)

%%% Local Variables:
%%% mode: latex
%%% TeX-master: strucfunc_paper
%%% End: