### 2022-12-20 table-fit-parameters.tex
Javier changes before submission

\begin{table*}
\begin{center}\caption{Fitting parameters.}
\begin{tabular}{cCCCCCCCC}\toprule
\hii{}    &  \text{No. points} & \text{Relatively}  & \text{Weight / No. points}     &  \text{Weight / No. points}     & \text{Max.}  & \text{Reduced} & \text{Autocorrelation } & \text{Acceptance}\\
Region    &  \text{merged} & \text{uncertainty} & \text{Small scales} &  \text{Large scales} & \text{separation}  & \text{chi-square} & \text{time} & \text{fraction}\\
\midrule
Orion     & 3 & 0.02    &  -           &   -        & 0.5L  & 0.82 & 60  & 0.55\\
EON       & - & 0.03   &  2.0 / 3     &   3.0 / 3        & 0.7L  & 0.62 & 40 & 0.64\\
Lagoon    & 2 & 0.08    &  2.5 / 16    &   2.5 / 8  & 0.5L  & 0.94 & 72  & 0.50\\
Carina    & 2 & 0.055   &  3.0 / 14    &   3.0 / 4  & 0.5L  & 0.94 & 72  & 0.50\\
30 Dor    & 5 &  0.07   &  -           &   -        & 0.9L  & 0.98 & 84  & 0.54\\
NGC 346   & 3 & 0.02    &  -           &   -        & 0.5L  & 0.83 & 64  & 0.54\\
Hubble V  & - & 0.05    &  2.0 / 3     &   -        & 0.6L  & 0.86 & 340 & 0.44\\
Hubble X  & - & 0.032   &  2.0 / 4     &   2.0 / 6  & 0.5L  & 0.73 & 79  & 0.49\\
NGC 595   & - &  0.055  &  4.5 / 3     &   3.0 / 5  & 0.5L  & 0.83 & 67  & 0.51\\
NGC 604   & - &  0.08  &   2.0 / 3    &   1.5 / 2 & 0.5L  & 0.92 & 112 & 0.45\\
\bottomrule
\end{tabular}\label{tab:fitting-parameters}
\end{center}
\end{table*}