### 2023-04-21 reply-to-referee.txt
Write the reply to referee

			 ______________________

			    REPLY TO REFEREE

			  García-Vázquez et al
			 ______________________

,----
| will be happy to recommend the manuscript for publication after some
| clarification of some details:
| 
| To study the Finite box effects, as well as that of seeing the authors
| do an interesting exercise with idealized data (as fractional Brownian
| motion -fBm- 2D data modified to include a correlation taper).  This
| approach is very clear, but there is one issue that requires
| attention, and that is that of the projection from a 3D field to a 2D
| field.  In reality the density/velocity fields are three-dimensional
| but some information is lost in the projection to a 2D map.  For the
| exercise presented in appendix A, the authors use a 2D map that was
| modified to include the proper spectral index (n_2D = m +2) and a
| correlation length r_0, in reality one would have a 3D map with a
| spectral index (n_3D = m +3), which if integrated along the line of
| sight would change the spectral index from n_3D to n_2D (and within
| the transition some inertial range could be 'lost') .
| 
| In other words in the current version of the paper the r_0 is entirely
| in the plane of the sky, while in reality should correspond to the 3D
| field, and at the same time the transition from a 3D to a 2D map could
| curve more the structure functions.  I believe that it would be
| beneficial to obtain the 2D map from collapsing a 3D fBm (with an
| index m+3) and the same exponential taper used and verify that the
| main results hold, or discuss this an additional limitation.
`----

We thank the referee for the insightful comments. It is quite true that
our approach of directly simulating 2d maps is not strictly correct and
that we should really simulate three dimensional fields and then project
them down to two dimensions for further analysis. So, we have done this
and the results of Appendix A have now been completely replaced. It
turns out that the new results are extremely similar to the old ones, so
the changes to the text (highlighted in red in the manuscript) are
minimal.

One new issue that we had to grapple with in order to use this more
realistic approach is that we needed to simulate the three dimensional
emissivity field (assumed proportional to ionized density squared for H
alpha) as well as the three dimensional velocity field.  For this, we
used the results shown in our Figure 12 to estimate the variance in the
emissivity, which we assumed to follow a log-normal distribution.  For
the spatial properties of the emissivity fluctuations, we assumed that
the correlation length and power law index was the same as for the
velocity and that the velocity and emissivity fluctuations have no
mutual correlation.  We investigated three different values of the rms
emissivity dispersion: sigma_E = 0, 1, and 2 (in units of the mean
value). The results were very similar in all cases, and we only show the
sigma_E = 1 case in the paper, since this is the most appropriate for
the majority of our regions.

With the new results, we found an improved analytic approximation for
the effects of seeing (equation 15 and A1). We have therefore re-run all
of the model fits in section 4, but the results change by less than 1%.

We have also taken the opportunity to improve some of the descriptions
of individual regions in Appendix C.